---
title: "Plot PCA GGR"
output: html_notebook
---

Packages:
```{r setup}
library("knitr")
wd ="/mnt/archgen/HistoGenes/analysis/PCA_211018/"
opts_knit$set(root.dir = wd)
library(ggplot2)
```

Set the variables. 
+ fn1=ofn from the pca 
+ fn2=fn for the sampleinfo.csv and ofn of the plotting
```{r}
fn1 = "PCA_211018"  ## File prefix
fn2 = "PCA_211018_batch1_25"  ## File prefix of sampleinfo and ofn
PCnum = 7  ## # of top PCs to plot
tag= "_ewPlot_Seminar" ##additional tag for the output, not required
```


The working directory MUST contain: 
* The _poplist.txt file, a file that has a number of columns equal to the number of PCAs performed, each column containing the list of the pops used to calculate the PCs
* The _popinfo.csv file, containing all aliases for pops
* The _sampleinfo.csv file

POPINFO GUIDE: 
Columns to edit: Mark,Mark.sub
--> Mark.sub currentl not working :(
Codes: 0,0 -> dot no tag
       0,1 -> dot no tag
       1,0 -> dot and tag
       1,1 -> dot and tag

Examples for all these files are in `/home/luca_traverso/scripts/Rscripts/plotpca_exmls/` (unupdated)
```{r}
setwd(wd)
pop.list.file = paste(fn1, "_poplist.txt", sep="")
pop.info.file = paste(fn1, "_popinfo.csv", sep="")
sample.info.file = paste(fn2, "_sampleinfo.csv", sep="")
```

Import the data
```{r}
## Import a list of populations used for generating PCs
pop.list = as.matrix(read.table(pop.list.file, header=F,sep = ","))

## Import a list of ancient samples to be renamed/regrouped and to be plotted
sample.info = read.csv(sample.info.file, header=T,sep=",")

## import population code and generate random colors for each language family
popdata = read.csv(pop.info.file, header=T, sep = ",")
pops = as.vector(popdata$Pop.new)
pnames = as.vector(popdata$Pop.new)
popcode = as.vector(popdata$Symbol)
popLFs = as.vector(popdata$LF)

LFnames = sort(unique(popLFs))
set.seed(6602)
rgbcolvec = matrix(sample(0:255, length(LFnames)*3, replace=T), ncol=3, byrow=T)
LFcols = rgb(rgbcolvec[,1], rgbcolvec[,2], rgbcolvec[,3], maxColorValue=255)
evals = read.table(gzfile(paste(fn1, ".eval.txt.gz", sep="")), header=F) ## Import eigenvalue information
```



##Plotting

```{r}
for (l in 1:ncol(evals)) {
#if(l==3) {next}

num = l  ## repeat the entire script from 1 to ncol(pop.list)
print(l)

fname = paste(wd, fn1, "_", num, ".evec.txt.gz", sep="")

## Import data and change group labels for2 Avars to "PannonianAvar"
data = read.table(gzfile(fname), header=T)

## Update PID/IID of the listed samples
tsm.nums = c()
for (iid in as.vector(sample.info$IID)) tsm.nums = c(tsm.nums, which(as.vector(data$ID) == iid))

tpv1 = as.vector(data$pop); tpv1[tsm.nums] = as.vector(sample.info$PID.new)
tpv2 = as.vector(data$ID); tpv2[tsm.nums] = as.vector(sample.info$IID.new)
data = data.frame(tpv1, tpv2, data[,-1*(1:2)])
names(data)[1:2] = c("pop", "ID")
label.fv1 = tsm.nums[as.vector(sample.info$Mark) == 1]

pops.in = as.vector(as.matrix(pop.list[,num]))
pops.in = pops.in[!is.na(pops.in)]

## Mark individuals included in the calculation
## and generate individual vectors for color and two-letter codes
filvec = rep(F, nrow(data))
for (pn in 1:length(pops.in)) filvec[as.vector(data$pop) == pops.in[pn]] = T

popcodevec = rep(NA, nrow(data))
colvec = rep(NA, nrow(data))
for (pn in 1:length(pops)) {
  tnum1 = which(LFnames == popLFs[pn])
  popcodevec[as.vector(data$pop) == pops[pn]] = popcode[pn]
  colvec[as.vector(data$pop) == pops[pn]] = LFcols[tnum1]
}



## Generate a legend vector
lgv = c(); lgcv = c()
for (i in 1:length(pops)) {
  if (pops[i] %in% pops.in) {
    lgv = c(lgv, paste(popcode[i], pnames[i]))
    lgcv = c(lgcv, LFcols[LFnames == popLFs[i]])
  }
}

## Generate a legend vector for language groups (color code)

gr.lgv = c(); gr.lgcv = c()
for (i in 1:length(pops)) {
  if (pops[i] %in% pops.in) {
    if (!(popLFs[i] %in% gr.lgv)) {
      gr.lgv = c(gr.lgv, popLFs[i])
      gr.lgcv = c(gr.lgcv, LFcols[LFnames == popLFs[i]])
    }
  }
}
gr.lgv.ord = order(gr.lgv)
gr.lgcv = gr.lgcv[gr.lgv.ord]
gr.lgv = gr.lgv[gr.lgv.ord]








#Define ancient groups to be plotted
aps = unique(as.vector(sample.info$PID.new))
## aps = sort(aps)
als = gsub("Hovsgol", "Khovsgol", gsub(".SG", "", aps))


#Plotting
## rgbcolors = grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]
## set.seed(702203)
## acols = sample(rgbcolors, 200)
acols = c("red","green","lavender","brown", "yellow","blue","orange",
          "magenta","beige","darkseagreen3","deepskyblue3","pink",
          "maroon","purple", "cyan","darkolivegreen1",
          "pink", "aquamarine4", "aquamarine1", 
          "darkgoldenrod2", "coral", "navy", "grey", "black")
acols = rep(acols, 10)
apchs = rep(c(21:25),30)

a.fv = c(); a.cv = c(); a.pchv = c()
for (an in 1:length(aps)) {
  tfv1 = which(as.vector(data$pop) == aps[an])
  a.fv = c(a.fv, tfv1)
  a.cv = c(a.cv, rep(acols[an], length(tfv1)))
  a.pchv = c(a.pchv, rep(apchs[an], length(tfv1)))
}
filvec2 = c(which(filvec), a.fv)

pvs = as.vector(as.matrix(evals[,num]))
pvs = pvs[!is.na(pvs)]; pvs = pvs / sum(pvs) * 100
pvs2 = paste(round(pvs, digits=2), "%")


data$pc1=-data$pc1
if (l == 6) {
	data$pc2=-data$pc2
}
pdf.name = paste(fn2, "_", num, tag,".pdf", sep="")
pdf(pdf.name, height=14, width=13, colormodel="cmyk")
par(cex.lab=1.7, cex.axis=1.7, cex.main=2)
mm = cbind(c(1,3),c(2,3)); layout(mm, widths=c(9,4), heights=c(8.5,5.5))

for (i in 1:3) {
  for (j in (i+1):4) {
    x1 = min(data[filvec2, i+2]); x2 = max(data[filvec2, i+2]); xmin = x1 - (x2-x1)*0.05; xmax = x2 + (x2-x1)*0.05
    y1 = min(data[filvec2, j+2]); y2 = max(data[filvec2, j+2]); ymin = y1 - (y2-y1)*0.05; ymax = y2 + (y2-y1)*0.05

    xlabvec = paste("PC ", i, " (", pvs2[i], ")", sep="")
    ylabvec = paste("PC ", j, " (", pvs2[j], ")", sep="")

    par(mar=c(5.1, 5.1, 1.1, 1.1))
    plot(c(x1,x2), c(y1,y2), type="n", xlab=xlabvec, ylab=ylabvec, cex=1.5)
    text(x=data[filvec,i+2], y=data[filvec,j+2], labels=popcodevec[filvec], col=alpha(colvec[filvec],0.5), font=2, cex=0.8)
    points(data[a.fv,i+2], y=data[a.fv,j+2], pch=a.pchv, bg=a.cv, cex=1)
    text(x=data[label.fv1,i+2], y=data[label.fv1,j+2], label=as.vector(data$ID)[label.fv1], font=2, cex=0.2, pos=2)
    if (num == 5 & i == 1 & j == 2) {
      legend("bottomleft", title="Moderns by language family", legend=gr.lgv, text.col=gr.lgcv, ncol=3, bty="n", text.font=2, cex=1.2)
    }

    par(mar=c(5.1, 0.1, 1.1, 0.1))
    plot(c(0,1),c(0,1), type="n", xaxt="n", yaxt="n", xlab="", ylab="", bty="n")
    legend("topleft", title="Ancients", legend=als, pch=apchs, pt.bg=acols, text.font=2, cex=1.2, y.intersp=1.05, ncol=1, bty="n")

    plot(c(0,1),c(0,1), type="n", xaxt="n", yaxt="n", xlab="", ylab="", bty="n")
    legend("top", legend=lgv, text.col=lgcv, ncol=6, bty="n", text.font=2)

  }
}
dev.off()

pdf.name = paste(fn2, "_", num, tag,"_bg.pdf", sep="")
pdf(pdf.name, height=14, width=13, colormodel="cmyk")
par(cex.lab=1.7, cex.axis=1.7, cex.main=2)
mm = cbind(c(1,3),c(2,3)); layout(mm, widths=c(9,4), heights=c(8.5,5.5))

for (i in 1:3) {
  for (j in (i+1):4) {
    x1 = min(data[filvec2, i+2]); x2 = max(data[filvec2, i+2]); xmin = x1 - (x2-x1)*0.05; xmax = x2 + (x2-x1)*0.05
    y1 = min(data[filvec2, j+2]); y2 = max(data[filvec2, j+2]); ymin = y1 - (y2-y1)*0.05; ymax = y2 + (y2-y1)*0.05

    xlabvec = paste("PC ", i, " (", pvs2[i], ")", sep="")
    ylabvec = paste("PC ", j, " (", pvs2[j], ")", sep="")

    par(mar=c(5.1, 5.1, 1.1, 1.1))
    plot(c(x1,x2), c(y1,y2), type="n", xlab=xlabvec, ylab=ylabvec, cex=1.5)
    text(x=data[filvec,i+2], y=data[filvec,j+2], labels=popcodevec[filvec], col="grey80", font=2, cex=1.0)
    points(data[a.fv,i+2], y=data[a.fv,j+2], pch=a.pchv, bg=a.cv, cex=1)
    text(x=data[label.fv1,i+2], y=data[label.fv1,j+2], label=as.vector(data$ID)[label.fv1], font=2, cex=0.4, pos=2)

    par(mar=c(5.1, 0.1, 1.1, 0.1))
    plot(c(0,1),c(0,1), type="n", xaxt="n", yaxt="n", xlab="", ylab="", bty="n")
    #legend("topleft", title="tv=TransversionOnly", legend=als, pch=apchs, pt.bg=acols, text.font=2, cex=1.2, y.intersp=1.05, ncol=2, bty="n")    
    legend("topleft", title="", legend=als, pch=apchs, pt.bg=acols, text.font=2, cex=1.2, y.intersp=1.05, ncol=1, bty="n")

    plot(c(0,1),c(0,1), type="n", xaxt="n", yaxt="n", xlab="", ylab="", bty="n")
    legend("top", legend=lgv, text.col=lgcv, ncol=6, bty="n", text.font=2)
  }
}
dev.off()


}
```
